---
title: "Tutorial on how to retrieve data from the INBOVEG database"
author: "Hans Van Calster & Els De Bie"
date: "1 maart 2019 (updated `r Sys.Date()`)"
output: html_document
---


# Packages and connection

The following packages are needed to run this code:

* knitr
* tidyverse
* DBI
* odbc
* glue

Be sure you have reading-rights for CYDONIA
otherwise place an ICT-call (ict.helpdesk@inbo.be)


```{r}

library(tidyverse)
library(DBI)
library(glue)
library(knitr)
library(odbc)

opts_chunk$set(echo = TRUE)

```
The following R-code can be used to establish a connection to INBOVEG by means of a connection string:

<!--better to use a connection string than dsn. dsn requires extra steps and settings in windows odbc manager-->

```{r, eval = FALSE}
con <- dbConnect(odbc::odbc(), .connection_string = "Driver=SQL Server;Server=inbo-sql07-prd.inbo.be,1433;Database=D0010_00_Cydonia;Trusted_Connection=Yes;")
```


# Retrieving data

## *iv_Survey*:

gives the list of all surveys in InboVeg

- define the name of the survey by survey <- "name"

```{r}

survey_info <- function(survey, con) {
  dbGetQuery(con, glue_sql(
    "SELECT
    ivS.Id
    , ivS.Name
    , ivS.Description
    , ivS.Owner
    , ivS.creator
    FROM [dbo].[ivSurvey] ivS
    WHERE ivS.Name LIKE {survey}", 
    ivS.Name = survey,
    .con = con ))
}

```

Example
  survey <- "OudeLanden_1979"
  SurveyInfo <- survey_info(survey, con)
  SurveyInfo
  
The whole list of surveys is given by  
AllSurveys <- survey_info(survey = "%", .con = con)

Only a part of the survey name is known? Or all the surveys of "MILKLIM" for example want to be given
Part <-"torf"
PartSurvey <- survey_info(survey = (str_detect(tolower(Name), "PART"))) ## dit werkt dus nog niet!!


## *iv_headerinfo*: 

- contains metadata for a vegetation-relevé (one row per vegetation-relevé identified by 'RecordingGivid')
- specify two parameters for the function:
    - RecType = c('Classic', 'Classic-emmer', 'Classic-ketting', 'BioHab', 'ABS')
    - SurveyName = to get the list, run the code under "## iv_survey

RecType <- "Classic"
SurveyName <- "input_name_survey"
    
```{r}

header_info <- function(SurveyName, RecType, .con) {
  dbGetQuery(con, glue_sql(
    "SELECT 
      ivR.[RecordingGivid]
      , ivS.Name
      , ivR.UserReference
      , ivR.LocationCode
      , ivR.Latitude
      , ivR.Longitude
      , ivR.Area
      , ivR.Length
      , ivR.Width
      , ivR.SurveyId
      , ivR.RecTypeID
      , coalesce(area, convert( nvarchar(20),ivR.Length * ivR.Width)) as B
      FROM [dbo].[ivRecording] ivR
      INNER JOIN [dbo].[ivSurvey] ivS on ivS.Id = ivR.SurveyId
      INNER JOIN [dbo].[ivRecTypeD] ivRec on ivRec.ID = ivR.RecTypeID 
      where ivR.NeedsWork = 0
      AND ivS.Name LIKE {SurveyName}
      AND ivREc.Name LIKE {RecType}", 
    ivS.Name = SurveyName,
    ivRec.Name = RecType,
    .con = con))
}


```

Give up survey-name and Recordtype in this function

Headerinfo <- header_info(SurveyName, RecType, con)
Example
  Headerinfo <- header_info("OudeLanden_1979", "Classic", con)


## *iv_Classification_N2000*:

- gives the N2000code, recorderd by the observer of the relevé
- specify the name of the survey you want to use. if none, all the classification records in inboveg will be given
- specify the N2000 code to retrieve all relevés indicated as this code

SurveyName <- "input_name_survey"
N2000 <- "N2000code"

 
```{r}
classification_info <- function(SurveyName, N2000, .con) {
  dbGetQuery(con, glue_sql(
    "SELECT 
    ivR.RecordingGivid
    , ivS.Name as survey
    , ivRLClas.Classif
    --, ivRLClas.ClassifResource
    , ivRLRes_Class.ActionGroup
    , ivRLRes_Class.ListName
    --, ftAGL_Class.ListGIVID
    --, ftBWK.ListGIVID
    , ftBWK.Description as LocalClassification
    , ftN2k.Description  as Habitattype
    , ivRLClas.Cover
    , ftC.PctValue
    FROM ivRecording ivR
    INNER JOIN ivSurvey ivS on ivS.Id = ivR.surveyId
    LEFT JOIN [dbo].[ivRLClassification] ivRLClas on ivRLClas.RecordingID = ivR.Id
    LEFT JOIN [dbo].[ivRLResources] ivRLRes_Class on ivRLRes_Class.ResourceGIVID = ivRLClas.ClassifResource
    LEFT JOIN [syno].[Futon_dbo_ftActionGroupList] ftAGL_Class on ftAGL_Class.ActionGroup = ivRLRes_Class.ActionGroup collate Latin1_General_CI_AI
    AND ftAGL_Class.ListName = ivRLRes_Class.ListName collate Latin1_General_CI_AI
    --MAAR hoe los ik dit op als er twee type lijsten aanhangen? gevolg van vroeger opdeling in Local en AnnexI classification?
    LEFT JOIN [syno].[Futon_dbo_ftBWKValues] ftBWK on ftBWK.Code = ivRLClas.Classif collate Latin1_General_CI_AI 
    AND ftBWK.ListGIVID = ftAGL_Class.ListGIVID 
    LEFT JOIN [syno].[Futon_dbo_ftN2kValues] ftN2K on ftN2K.Code = ivRLClas.Classif collate Latin1_General_CI_AI 
    AND ftN2K.ListGIVID = ftAGL_Class.ListGIVID 
    LEFT JOIN [dbo].[ivRLResources] ivRLR_C on ivRLR_C.ResourceGIVID = ivRLClas.CoverResource
    LEFT JOIN [syno].[Futon_dbo_ftActionGroupList] ftAGL_C on ftAGL_C.ActionGroup = ivRLR_C.ActionGroup collate Latin1_General_CI_AI
    AND ftAGL_C.ListName = ivRLR_C.ListName collate Latin1_General_CI_AI
    LEFT JOIN [syno].[Futon_dbo_ftCoverValues] ftC on ftC.Code = ivRLClas.Cover collate Latin1_General_CI_AI
    AND ftAGL_C.ListGIVID = ftC.ListGIVID 
    WHERE ivRLClas.Classif is not NULL 
    AND ivS.Name LIKE {SurveyName}
    AND ivRLClas.Classif LIKE {N2000}",
           ivS.Name = SurveyName,
           ivRLClas.Classif = N2000, 
           .con = con))
}

```

Classifiction <- classification_info("OudeLanden_1979", "N2000", con)

## iv_Classification_BWK

- gives the BWK-code, recorderd by the observer of the relevé
- specify the name of the survey you want to use. if none, all the classification records in inboveg will be given
- specify the bwk-code to retrieve all relevés indicated as this code

SurveyName <- "input_name_survey"
BWK <- "bwkcode"

 
```{r}
classification_info <- function(SurveyName, BWK, .con) {
  dbGetQuery(con, glue_sql(
    "SELECT 
    ivR.RecordingGivid
    , ivS.Name as survey
    , ivRLClas.Classif
    , ivRLRes_Class.ActionGroup
    , ivRLRes_Class.ListName
    , ftBWK.Description as LocalClassification
    , ftN2k.Description  as Habitattype
    , ivRLClas.Cover
    , ftC.PctValue
    FROM ivRecording ivR
    INNER JOIN ivSurvey ivS on ivS.Id = ivR.surveyId
    LEFT JOIN [dbo].[ivRLClassification] ivRLClas on ivRLClas.RecordingID = ivR.Id
    LEFT JOIN [dbo].[ivRLResources] ivRLRes_Class on ivRLRes_Class.ResourceGIVID = ivRLClas.ClassifResource
    LEFT JOIN [syno].[Futon_dbo_ftActionGroupList] ftAGL_Class on ftAGL_Class.ActionGroup = ivRLRes_Class.ActionGroup collate Latin1_General_CI_AI
    AND ftAGL_Class.ListName = ivRLRes_Class.ListName collate Latin1_General_CI_AI
    --MAAR hoe los ik dit op als er twee type lijsten aanhangen? gevolg van vroeger opdeling in Local en AnnexI classification?
    LEFT JOIN [syno].[Futon_dbo_ftBWKValues] ftBWK on ftBWK.Code = ivRLClas.Classif collate Latin1_General_CI_AI 
    AND ftBWK.ListGIVID = ftAGL_Class.ListGIVID 
    LEFT JOIN [syno].[Futon_dbo_ftN2kValues] ftN2K on ftN2K.Code = ivRLClas.Classif collate Latin1_General_CI_AI 
    AND ftN2K.ListGIVID = ftAGL_Class.ListGIVID 
    LEFT JOIN [dbo].[ivRLResources] ivRLR_C on ivRLR_C.ResourceGIVID = ivRLClas.CoverResource
    LEFT JOIN [syno].[Futon_dbo_ftActionGroupList] ftAGL_C on ftAGL_C.ActionGroup = ivRLR_C.ActionGroup collate Latin1_General_CI_AI
    AND ftAGL_C.ListName = ivRLR_C.ListName collate Latin1_General_CI_AI
    LEFT JOIN [syno].[Futon_dbo_ftCoverValues] ftC on ftC.Code = ivRLClas.Cover collate Latin1_General_CI_AI
    AND ftAGL_C.ListGIVID = ftC.ListGIVID 
    WHERE ivRLClas.Classif is not NULL 
    AND ivS.Name LIKE {SurveyName}
    AND ivRLClas.Classif LIKE {BWK}",
           ivS.Name = SurveyName,
           ivRLClas.Classif = BWK, 
           .con = con))
}

```

Classifiction <- classification_info("OudeLanden_1979", "BWK", con)

## iv_Relevés

Getting the relevés of one Survey

```{r, include = FALSE}

Vegetation_info <- function(SurveyName, .con) {
   dbGetQuery(con, glue_sql(
            "SELECT 
            ivR.[RecordingGivid]
        	, ivRL_Layer.LayerCode
        	, ivRL_Layer.CoverCode
        	, ivRL_Iden.TaxonFullText as OrignalName
        	, Synoniem.ScientificName
        	, ivRL_Iden.PhenologyCode
        	, ivRL_Taxon.CoverageCode
        	, ftCover.PctValue
        	, ftAGL.Description as RecordingScale
        FROM [dbo].[ivRecording] ivR 
  -- Deel met soortenlijst en synoniem
        INNER JOIN [dbo].[ivRLLayer] ivRL_Layer on ivRL_Layer.RecordingID = ivR.Id
        INNER JOIN [dbo].[ivRLTaxonOccurrence] ivRL_Taxon on ivRL_Taxon.LayerID = ivRL_Layer.ID
        INNER JOIN [dbo].[ivRLIdentification] ivRL_Iden on ivRL_Iden.OccurrenceID = ivRL_Taxon.ID
        LEFT JOIN (SELECT ftTaxon.TaxonName AS TaxonFullText
        			 , COALESCE([GetSyn].TaxonName, ftTaxon.TaxonName) AS ScientificName
        			 , COALESCE([GetSyn].TaxonGIVID, ftTaxon.TaxonGIVID) AS TAXON_LIST_ITEM_KEY
        			 , COALESCE([GetSyn].TaxonQuickCode, ftTaxon.TaxonQuickCode) AS QuickCode
        		  FROM [syno].[Futon_dbo_ftTaxon] ftTaxon
        		  INNER JOIN [syno].[Futon_dbo_ftTaxonListItem] ftTLI ON ftTLI.TaxonGIVID = ftTaxon.TaxonGIVID 
        	      LEFT JOIN (SELECT ftTaxonLI.TaxonListItemGIVID
        						, ftTaxon.TaxonGIVID
        						, ftTaxon.TaxonName
        						, ftTaxon.TaxonQuickCode
        						, ftAGL.ListName
        						, ftTaxonLI.PreferedListItemGIVID
        					FROM [syno].[Futon_dbo_ftActionGroupList] ftAGL 
        					INNER JOIN [syno].[Futon_dbo_ftTaxonListItem] ftTaxonLI ON ftTaxonLI.TaxonListGIVID = ftAGL.ListGIVID 
        					LEFT JOIN [syno].[Futon_dbo_ftTaxon] ftTaxon ON ftTaxon.TaxonGIVID = ftTaxonLI.TaxonGIVID 
        					WHERE 1=1
        					AND ftAGL.ListName = 'INBO-2011 Sci'	
        					) GetSyn ON GetSyn.TaxonListItemGIVID = ftTLI.PreferedListItemGIVID
        		WHERE ftTLI.TaxonListGIVID = 'TL2011092815101010'
        		) Synoniem on ivRL_Iden.TaxonFullText = Synoniem.TaxonFullText collate Latin1_General_CI_AI
        		--ORDER BY ivR.RecordingGIVID
  --Hier begint deel met bedekking
        LEFT JOIN [dbo].[ivRLResources] ivRL_Res on ivRL_Res.ResourceGIVID = ivRL_Taxon.CoverageResource
        LEFT JOIN [syno].[Futon_dbo_ftActionGroupList] ftAGL on ftAGL.ActionGroup = ivRL_Res.ActionGroup collate Latin1_General_CI_AI
        		AND ftAGL.ListName = ivRL_Res.ListName collate Latin1_General_CI_AI
        LEFT JOIN [syno].[Futon_dbo_ftCoverValues] ftCover on ftCover.ListGIVID = ftAGL.ListGIVID
        		AND ivRL_Taxon.CoverageCode = ftCover.Code collate Latin1_General_CI_AI
        WHERE ivR.NeedsWork = 0
            AND ivRL_Iden.Preferred = 1
            AND ivS.Name LIKE {SurveyName}", 
               ivS.Name = SurveyName,
               .con = con ))
}

```

Vegetatieopname werkt nog niet, hoe omgaan met 2de databank bevragen. hier zoals in Ssms (syno) gewerkt, de tabellen zijn toegevoegd door joL aan Cydonia conncetie via die Syno....

To close the connection:

```{r, eval = FALSE}
dbDisconnect(con)

rm(con)
```

